05_DESeq_manipulation

### 2019-05-14 analysis/05_DESeq2_manipulate.Rmd
with better order
now with a heatmap of top 30 DEGs

m.colData$outcome <- ifelse(grepl("d3|d9|d17|n2", m.colData$treatment), "remove", 
                     ifelse(grepl("d8", m.colData$treatment),"early",
                     ifelse(grepl("prolong", m.colData$treatment),"prolong",
                     ifelse(grepl("extend", m.colData$treatment),"extend", NA))))
m.colData$outcome <- factor(m.colData$outcome, levels = 
                              c("remove",  "prolong",
                                "early", "extend"))
summary(m.colData[c(7,3,4,5,9)])
```

@@

```{r althings, eval = FALSE}

@@

#pheatmap(sampleDistMatrix,
#         clustering_distance_rows=sampleDists,
#         clustering_distance_cols=sampleDists,
#         col=colors,
 #        fontsize = 6, 
 #        main = eachgroup)
}
```

```{r allthings-outcome}

@@

allcontrasts <- dat %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#FFFFFF00") + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

@@

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
#print(pca34)
}
```

>>>>

m.colData$outcome <- ifelse(grepl("d3|d9|d17|n2", m.colData$treatment), "remove", 
                     ifelse(grepl("d8", m.colData$treatment),"early",
                     ifelse(grepl("prolong", m.colData$treatment),"prolong",
                     ifelse(grepl("extend", m.colData$treatment),"extend", NA))))
m.colData$outcome <- factor(m.colData$outcome, levels = 
                              c("remove",  "prolong",
                                "early", "extend"))
summary(m.colData[c(7,3,4,5,8,9)])
```

@@

```{r althings, eval = T}

## subset for test

m.colData <- m.colData %>%
  filter(sextissue == "female_pituitary") %>%
  droplevels()

@@

#pheatmap(sampleDistMatrix,
#         clustering_distance_rows=sampleDists,
#         clustering_distance_cols=sampleDists,
#         col=colors,
 #        fontsize = 6, 
 #        main = eachgroup)

# Multi-factor designs
# revals is my function easily calculate multiple multi factor designs
# it is built around the documentaion from DESeq which looks more simply like
# res <- results(dds,contrast=c("columnname", "factor1", "factor2"))
# then I extract the p-values from within the results for showing number of DEGes
resvals <- function(contrastvector, mypval){
  res <- results(dds, contrast = c(contrastvector[1],contrastvector[2],contrastvector[3]), independentFiltering = T)
  sumpvalue <- sum(res$pvalue < mypval, na.rm = TRUE)
  #print(sumpvalue)
  sumpadj <- sum(res$padj < mypval, na.rm = TRUE)
  print(sumpadj)
  vals <- cbind(res$pvalue, res$padj)
  pvalcolname <- as.character(paste("pval",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
  padjcolname <- as.character(paste("padj",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
  colnames(vals) <- c(pvalcolname, padjcolname)
  return(vals)
}
## DEG by contrasts at 0.1 pvalue
contrast1 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d9'), mypval = 0.01) #12
contrast2 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d17'), mypval = 0.01) #73
contrast3 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.n2'), mypval = 0.01) #284
contrast4 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d8'), mypval = 0.01) #15
contrast5 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'prolong'), mypval = 0.01) #555
contrast6 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'extend'), mypval = 0.01) #1109
contrast7 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.inc.d17'), mypval = 0.01) #280
contrast8 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.n2'), mypval = 0.01) #440
contrast9 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.inc.d8'), mypval = 0.01) #484
contrast10 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'prolong'), mypval = 0.01) #529
contrast11 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'extend'), mypval = 0.01) #1177

contrast12 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'm.n2'), mypval = 0.01) #5
contrast13 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'm.inc.d8'), mypval = 0.01) #553
contrast14 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'prolong'), mypval = 0.01) #339
contrast15 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'extend'), mypval = 0.01) #782
contrast16 <- resvals(contrastvector = c('treatment', 'm.n2', 'm.inc.d8'), mypval = 0.01) #1123
contrast17 <- resvals(contrastvector = c('treatment', 'm.n2', 'prolong'), mypval = 0.01) #265
contrast18 <- resvals(contrastvector = c('treatment', 'm.n2', 'extend'), mypval = 0.01) #552
contrast19 <- resvals(contrastvector = c('treatment', 'm.inc.d8', 'prolong'), mypval = 0.01) #828
contrast20 <- resvals(contrastvector = c('treatment', 'm.inc.d8', 'extend'), mypval = 0.01) #1110
contrast21 <- resvals(contrastvector = c('treatment', 'prolong', 'extend'), mypval = 0.01) #64

DEGs <- assay(vsd)
DEGs <- cbind(DEGs, contrast1, contrast2, contrast3, contrast4, contrast5,
              contrast6, contrast7, contrast8, contrast9, contrast10,
              contrast11, contrast12, contrast13, contrast14, contrast15,
              contrast16, contrast17, contrast18, contrast19, contrast20,contrast21)
DEGs <- as.data.frame(DEGs) # convert matrix to dataframe
DEGs$rownames <- rownames(DEGs)  # add the rownames to the dataframe

#DEGs[is.na(DEGs)] <-  1

DEGs$padjmin <- with(DEGs, pmin(pvaltreatmentm.inc.d3m.inc.d9, 
  padjtreatmentm.inc.d3m.inc.d9,               
  pvaltreatmentm.inc.d3m.inc.d17, padjtreatmentm.inc.d3m.inc.d17,              
  pvaltreatmentm.inc.d3m.n2, padjtreatmentm.inc.d3m.n2,                  
  pvaltreatmentm.inc.d3m.inc.d8, padjtreatmentm.inc.d3m.inc.d8,               
  pvaltreatmentm.inc.d3prolong, padjtreatmentm.inc.d3prolong,               
  pvaltreatmentm.inc.d3extend, padjtreatmentm.inc.d3extend,                
  pvaltreatmentm.inc.d9m.inc.d17, padjtreatmentm.inc.d9m.inc.d17,             
  pvaltreatmentm.inc.d9m.n2, padjtreatmentm.inc.d9m.n2,                  
  pvaltreatmentm.inc.d9m.inc.d8, padjtreatmentm.inc.d9m.inc.d8,              
  pvaltreatmentm.inc.d9prolong, padjtreatmentm.inc.d9prolong,               
  pvaltreatmentm.inc.d9extend, padjtreatmentm.inc.d9extend,                
  pvaltreatmentm.inc.d17m.n2, padjtreatmentm.inc.d17m.n2,                 
  pvaltreatmentm.inc.d17m.inc.d8, padjtreatmentm.inc.d17m.inc.d8,             
  pvaltreatmentm.inc.d17prolong, padjtreatmentm.inc.d17prolong,
  pvaltreatmentm.inc.d17extend,padjtreatmentm.inc.d17extend, 
  pvaltreatmentm.n2m.inc.d8, padjtreatmentm.n2m.inc.d8,                  
  pvaltreatmentm.n2prolong, padjtreatmentm.n2prolong,                   
  pvaltreatmentm.n2extend, padjtreatmentm.n2extend,                    
  pvaltreatmentm.inc.d8prolong, padjtreatmentm.inc.d8prolong,               
  pvaltreatmentm.inc.d8extend, padjtreatmentm.inc.d8extend,                
  pvaltreatmentprolongextend, padjtreatmentprolongextend))

sigDEGs <- DEGs %>% arrange(padjmin)
sigDEGs <- head(sigDEGs,30)
sigDEGs <- as.data.frame(sigDEGs)
rownames(sigDEGs) <- sigDEGs$rownames
drop.cols <-colnames(sigDEGs[,grep("padj|pval|pmin|rownames", colnames(sigDEGs))])
sigDEGs <- sigDEGs %>% dplyr::select(-one_of(drop.cols))
sigDEGs <- as.matrix(sigDEGs)
sigDEGs <- sigDEGs - rowMeans(sigDEGs)

paletteLength <- 30
myBreaks <- c(seq(min(sigDEGs), 0, length.out=ceiling(paletteLength/2) + 1), 
                seq(max(sigDEGs)/paletteLength, max(sigDEGs), length.out=floor(paletteLength/2)))

anndf <- m.colData %>% dplyr::select(treatment)
rownames(anndf) <- m.colData$V1

sigDEGs <- as.matrix(sigDEGs) 
pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf)

pheatmap(sigDEGs, kmeans_k = 5,
         show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf)
}
```

```{r allthings-outcome, eval = F}

@@

@@

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
#print(pca34)

}
```
